Bayesian RKHS-based methods in functional regression

### 2022-06-08 slides/seio.tex
Changes in slides

\begin{frame}{RKHS reminder}
  \begin{definition}
    Let \(K\) be continuous, and define the preliminary space
    \[
    \Hcal_0(K) = \left\{ f \in L^2[0,1]: \ f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot),  p \in \N,  a_i \in \R,  t_i \in [0, 1] \right\},
    \]
    which is endowed with the scalar product \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\), for \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\). Then, \(\Hcal(K)\) is the completion of \(\Hcal_0(K)\) under this scalar product.
  \end{definition}

@@

  Note that \(\Hcal(K)\subset L^2[0, 1]\), but the elements of \(\Hcal(K)\) are simpler and more manageable in general. Plus, the latter is a space of \maroon{functions}, and not of equivalence classes.
\end{frame}

@@

\begin{frame}{A new functional parameter...}
  \begin{itemize}
    \item The initial idea was to suppose \(\beta \in \Hcal(K)\).
    \item To further simplify the model, we will assume \(\beta \in \Hcal_0(K)\), which is dense in \(\Hcal(K)\) by construction. That is,
    \[
      \beta(\cdot) = \sum_{j=1}^p \beta_j K(t_j, \cdot).
    \]
    \item When evaluating \(\dotprod{x}{\beta}_K\) for a trajectory \(x=X(\omega)\), we resort to Loève's isometry:
    \[
      \dotprod{x}{\beta}_K = \Psi_x(\beta) = \sum_{j=1}^p \beta_j X(t_j).
    \]
  \end{itemize}

@@

\begin{frame}{...and two new models}
  Given a data set \(\mathcal D_n = \{(X_i, Y_i): i=1,\dots, n\}\) composed of i.i.d. samples of \((X, Y)\), we consider:

@@

\begin{frame}{Posterior approximation}
  % We use the Python library \maroon{emcee}, which implements an MCMC algorithm consisting on multiple chains that interact with each other to advance on each step (\textit{Affine-invariant ensemble sampler}).

  \begin{figure}
    \includegraphics[width=.9\textwidth]{img/posterior}
    \caption{Example of an approximated posterior distribution \(\theta|\mathcal D_n\).}
  \end{figure}
\end{frame}

@@

@@

    \begin{figure}
      \includegraphics[width=\textwidth]{img/tau_posterior}
      \caption{Approximate posterior distribution for the impact points.}
    \end{figure}
\end{frame}

@@

\begin{frame}{Predictions (cont.)}
    At any rate, all the above predictors can be obtained after a single run of the MCMC algorithm.
  \begin{figure}
    \includegraphics[width=.75\textwidth]{img/ppc_lin_standalone}
    \caption{Example of \textit{posterior predictive distribution} in linear regression.}
  \end{figure}
\end{frame}

>>>>

\begin{frame}{RKHS reminder}
  \begin{definition}
    Let the covariance function \(K\) be continuous, and define the preliminary space
    \[
    \Hcal_0(K) = \left\{ f \in L^2[0,1]: \ f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot),  p \in \N,  a_i \in \R,  t_i \in [0, 1] \right\},
    \]
    which is endowed with the scalar product \(\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j)\), for \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\). Then, \(\Hcal(K)\) is the completion of \(\Hcal_0(K)\) under this scalar product.
  \end{definition}

@@

  Note that \(\Hcal(K)\subset L^2[0, 1]\) as subsets, but the elements of \(\Hcal(K)\) are simpler and more manageable in general. Plus, the latter is a space of \maroon{functions}, and not of equivalence classes.
\end{frame}

@@

\begin{frame}{A new functional parameter...}
  \begin{itemize}
    \item The initial idea was to suppose \(\beta \in \Hcal(K)\).
    \item To further simplify the model, we will assume \(\beta \in \Hcal_0(K)\), which is dense in \(\Hcal(K)\) by construction. That is,
    \[
      \beta(\cdot) = \sum_{j=1}^p \beta_j K(t_j, \cdot), \quad \beta_j \in \R, \ t_j \in [0, 1], \ p \in \N.
    \]
    \item When evaluating \(\dotprod{x}{\beta}_K\) for a trajectory \(x=X(\omega)\), we resort to Loève's isometry:
    \[
      \dotprod{x}{\beta}_K = \Psi_x(\beta) = \sum_{j=1}^p \beta_j X(t_j).
    \]
  \end{itemize}

@@

\begin{frame}{...and two new models}
  Given a data set \(\mathcal D_n = \{(X_i, Y_i): i=1,\dots, n\}\) composed of i.i.d. samples of \((X, Y)\), for \(\beta \in \Hcal_0(K)\) we consider:

@@

% \begin{frame}{Posterior approximation}
%  We use the Python library \maroon{emcee}, which implements an MCMC algorithm consisting on multiple chains that interact with each other to advance on each step (\textit{Affine-invariant ensemble sampler}).
%
%   \vspace{1em}
%
%   \begin{figure}
%     \includegraphics[width=.9\textwidth]{img/posterior}
%     \caption{Example of an approximated posterior distribution \(\theta|\mathcal D_n\).}
%   \end{figure}
% \end{frame}

@@

\begin{frame}{Predictions (cont.)}
    \vspace{1em}
  \begin{figure}
    \includegraphics[width=.75\textwidth]{img/ppc_lin_standalone}
    \caption{Example of \textit{posterior predictive distribution} in linear regression.}
  \end{figure}
\end{frame}

@@

    \begin{figure}
      \includegraphics[width=\textwidth]{img/tau_posterior}
      \caption{Approximate posterior distribution for the impact points.}
    \end{figure}

    At any rate, all the above predictors can be obtained after a single run of the MCMC algorithm.
\end{frame}

@@
